Pillar 1 - Sequential Pattern Mining

### 2019-01-24 Pillar 1 - Sequential Pattern Mining.Rmd
add save file

manual_confidence_rules <- pass_rules_df %>%
  separate(LHS, into=c("LHS Course ID", "LHS Score"), sep="_", remove= FALSE) %>%
  separate(RHS, into=c("RHS Course ID", "RHS Score"), sep="_", remove= FALSE) %>%
  mutate(`RHS Course ID`= substr(`RHS Course ID`, start=4, stop=10))%>%
  left_join(lookup_table, by = c("RHS Course ID"="Course ID")) %>%

@@

  # Filter and edit output
  filter(str_detect(`RHS Score`, pattern = "low"),
         str_detect(`LHS Score`, pattern = "low"),
         n_rule >= 5,
         Conf_corr > 0.2) %>%
  select(LHS, RHS, support, n_rule, Prob_both, n_both, Prop_low, Conf_corr, lift_corr, lift_diff) %>%
  mutate(support = round(support,3),
         Prob_both =round(Prob_both,3),
         Prop_low =round(Prop_low,3),
         Conf_corr= round(Conf_corr,3),
         lift_diff= round(lift_diff,3),
         lift_corr= round(lift_corr,3)) %>%
  arrange(lift_diff)
#Inspect results
manual_confidence_rules
```
##Arrange data for plots:
```{r}
for_plots <-  manual_confidence_rules %>%
  group_by(LHS)%>%
  mutate(LHS_count=n())
```
#Visualizing rules
```{r}
#look at counts
ggplot(for_plots, aes(LHS))+
  geom_bar()

@@

#Puting students into take and not take.

```{r}
test <- expand.grid(`Student ID`= unique(d_transcript$`Student ID`),`Course ID`= unique(d_course$`Course ID`)) %>%
  left_join(d_transcript, by=c("Course ID", "Student ID")) %>%
  select(`Course ID`, `Student ID`, Grade, Year_numerical, Period) %>%
  mutate(
    take_course = case_when(
      is.na(Grade)  ~ FALSE,
      T             ~ TRUE
      ),
    PFN = case_when(
      is.na(Grade)  ~ "not",
      Grade >= 5.5  ~ "pass",
      Grade < 5.5   ~ "fail"
      )
    )

@@

We want to change the calculation of confidence and support. The new rules are:
confidence: number of transactions of RHS/ number of transactions of RHS and compliment= p(A^B)/P(BUBc)
support:number of transactions of right hand side/number of transactions of RHS and compliment= p(A^B)/p(AUBUBc)
lift:

```{r}
pass_rules_df
test <- pass_rules_df %>%
  separate(LHS, into=c("c1","c2", "c3", "c4", "c5", "c6","c7", "c8", "c9", "c10"), sep= ",", remove=F) #has values until c6

>>>>

sequence_rules <- pass_rules_df %>%
  separate(LHS, into=c("LHS Course ID", "LHS Score"), sep="_", remove= FALSE) %>%
  separate(RHS, into=c("RHS Course ID", "RHS Score"), sep="_", remove= FALSE) %>%
  mutate(`RHS Course ID`= substr(`RHS Course ID`, start=4, stop=10))%>%
  left_join(lookup_table, by = c("RHS Course ID"="Course ID")) %>%

@@

  # Filter and edit output
  filter(str_detect(`RHS Score`, pattern = "low"),
         str_detect(`LHS Score`, pattern = "low"),
         n_rule >= 5,
         Conf_corr > 0.2) %>%
  select(LHS, RHS, support, n_rule, Prob_both, n_both, Prop_low, Conf_corr, lift_corr, lift_diff) %>%
  mutate(support = round(support,3),
         Prob_both =round(Prob_both,3),
         Prop_low =round(Prop_low,3),
         Conf_corr= round(Conf_corr,3),
         lift_diff= round(lift_diff,3),
         lift_corr= round(lift_corr,3)) %>%
  arrange(lift_diff)

```
##Save data
```{r}
save(sequence_rules,
     file = "sequence_rules.RDATA")
```

##Arrange data for plots:
```{r}
for_plots <-  manual_confidence_rules %>%
  group_by(LHS)%>%
  mutate(LHS_count=n())
```
#Visualizing rules
```{r}
#look at counts
ggplot(for_plots, aes(LHS))+
  geom_bar()

@@

#EXPERIMENTS:
##Condensing grades

```{r}

probabilities_manipulated <- d_transcript %>%
  select(-Course, -Year_numerical,-Period,-`Number Repeated Attempt`,-`Academic Year`)%>% #remove this before expanding grades
  mutate(Round_grade= round(Grade,0))
```

##Puting students into take and not take. 
```{r}

#this df has 731,676 rows:
test <- expand.grid(`Student ID`= unique(d_transcript$`Student ID`),`Course ID`= unique(d_course$`Course ID`)) %>%
  left_join(d_transcript, by=c("Course ID", "Student ID")) %>%
  select(`Course ID`, `Student ID`, Grade, Year_numerical, Period) %>%
  mutate(
    take_course = case_when(
      is.na(Grade)  ~ FALSE,
      T             ~ TRUE
      ),
    PFN = case_when(
      is.na(Grade)  ~ "not",
      Grade >= 5.5  ~ "pass",
      Grade < 5.5   ~ "fail"
      )
    )

@@
